---
title: "Patient data analysis"
author: "Tanaya Kavathekar"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    fig_height: 4.5
    fig_width: 7
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: true
  pdf_document:
    toc: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(error = F)
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = T)
knitr::opts_chunk$set(warning = F)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

```

```{r, include=F}
loadPkg("readxl")
loadPkg("plyr")
loadPkg("dplyr")
loadPkg("stringr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg("plotly")
loadPkg("lubridate")

```

Reading data ..
```{r}
#raw_df <- read_excel('master-data.xlsx',sheet='Sheet1')
raw_df <- read_excel('updated_GWG.xlsx',sheet='master')
raw_df2 <- read_excel('GWG.xlsx',sheet='master')

test_pt <- read_excel('test_patients_exclude.xlsx')
names(test_pt) <- str_replace_all(names(test_pt), c(" " = "_"))
length(unique(test_pt$Dim_Patient))

# if space in the column name change to underscore
names(raw_df2) <- str_replace_all(names(raw_df2), c(" " = "_"))
names(raw_df) <- str_replace_all(names(raw_df), c(" " = "_"))


dim(raw_df)
length(unique(raw_df$Dim_Patient))

length(unique(raw_df2$Dim_Patient))

```

Removing test patients 
```{r}

raw_df <- anti_join(raw_df, test_pt)
length(unique(raw_df$Dim_Patient))
```

Total number of rows and columns are `r dim(raw_df)` and total number of unique patients are `r length(unique(raw_df$Dim_Patient))`


Drop null values from the data - 
```{r}
df <- raw_df %>% drop_na(Dim_Patient)
df <- df %>% filter(Dim_Patient !=0)
dim(df)
length(unique(df$Dim_Patient))
df2 <- raw_df2 %>% drop_na(Observation_Value)
dim(df2)
length(unique(df2$Dim_Patient))
```

Remove duplicate rows:
```{r}

# remove diff column as it will be recalculated again
df <-subset(df,select = - c(Weight_Difference, Days_Since_Last_Entry))
# drop duplicates
df <- unique(df)

# GTA column has missing values 
sum(is.na(df))
# time being drop those columns 
df <-df %>% drop_na(GA_at_Observation)

#write_csv(df, "check_file.csv")

# extract week from the GTA_at_Obervation 
df$week <- sapply( df$GA_at_Observation,  function(x) 
  if(str_length(x) == 4) as.numeric(substr(x, 1,1)) 
  else if (str_length(x) == 5) as.numeric(substr(x, 1,2))
  else if (str_length(x) == 6) as.numeric(substr(x, 1,3))
  else if (str_length(x) == 7) as.numeric(substr(x, 1,4)))

length(unique(raw_df$Dim_Patient)) - length(unique(df$Dim_Patient))
length(unique(df$Dim_Patient))
print("total number of duplicate rows")
nrow(raw_df) - nrow(df)
```


Keep only those patients who have observation value > 80
```{r }
df <- df %>% filter(Observation_Value > 80) 
length(unique(df$Dim_Patient))
```

Keep only 47 weeks of data from observe date
```{r}

df <- df %>% filter((week>=0) &(week<=47))
# keep 322 days (46w) data from observ date 
#map <- df %>% group_by(Dim_Patient) %>% summarise(min_date =  min(Observed_At))
#map$Observed_till <- as.Date(map$min_date) + 322

#df <- merge(df, map[, c("Dim_Patient", "Observed_till")], by = "Dim_Patient", all.y = TRUE)

#df <- filter(df, df$Observed_At <= df$Observed_till)
length(unique(df$Dim_Patient))
```

Remove patients having only 3 or less observations and only one trimester data 
```{r}
# get only trimester values at patient-trimester level
dpt <- df %>% select(Dim_Patient, Trimester) %>% distinct(Dim_Patient, Trimester)
# check how many patients have length
agg <- aggregate(data=dpt, Trimester ~ Dim_Patient, function(x) length(unique(x)))

less_data <- agg %>% filter(Trimester == 1)

length(unique(less_data$Dim_Patient))

check2 <- merge(
  df, less_data, by = "Dim_Patient", all.y=T)

check2_2 <- check2 %>% group_by(Dim_Patient) %>% count()

# remove these patients 
ct2 <- check2_2 %>% filter(n<4)
length(unique(ct2$Dim_Patient))

df <- subset(df, !df$Dim_Patient %in% ct2$Dim_Patient)
length(unique(df$Dim_Patient))
```
Remove entries if less than 2 observations within a day
```{r}
intermediate_rollup <- df %>% group_by(Dim_Patient, date) %>% summarise(mean = mean(Observation_Value))
# remove entries if less than 2
count <- intermediate_rollup %>% group_by(Dim_Patient)  %>% count() %>% filter(n > 2)

df <- merge(df, count[,"Dim_Patient"], by = "Dim_Patient")
length(unique(df$Dim_Patient))
```

Day level check if value greater than 5
```{r}

df$date <- as.Date(df$Observed_At)

#p1 <- df %>% filter(Dim_Patient == "ref:Patient/Redox#advocateaurorahealth.org|1642714")
p1 <- df %>% filter(Dim_Patient == "ref:Patient/Redox#advocateaurorahealth.org|912816")

count_rows <- df %>% group_by(Dim_Patient) %>% summarise(initial_count = n())
  
daily <- p1 %>% group_by(Dim_Patient, date) %>% 
  mutate(weight_diff = c(0,diff(Observation_Value)), 
         wt_flg =  ifelse((abs(weight_diff) >= 5) & (Trimester != "PP"), NA, Observation_Value))  %>% ungroup() %>% filter(!is.na(wt_flg))


weekly <- daily %>% group_by(Dim_Patient, Trimester) %>% arrange(Observed_At) %>%
  mutate(weight_diff_tri = c(0,diff(Observation_Value)), 
         avg_wt_tri = getmode(Observation_Value),
         diff = Observation_Value - avg_wt_tri,
         wt_flg_tri =  ifelse((abs(diff) >= 15) & (Trimester != "PP"), NA, Observation_Value))%>% filter(!is.na(wt_flg_tri))%>% 
  ungroup() %>% 
  group_by(Dim_Patient) %>% 
  mutate(mode = getmode(Observation_Value), mode_dev =  ifelse((abs(Observation_Value - mode) >= 40), NA, Observation_Value)) %>% 
  filter(!is.na(mode_dev))


complete <- weekly %>% group_by(Dim_Patient, Trimester) %>% 
  mutate( avg = mean(Observation_Value)) %>% 
  ungroup() %>% 
  group_by(Dim_Patient) %>% 
  mutate(diff_avg =ifelse((abs(Observation_Value - avg[1]) >= 20) & (Trimester == "PP"), NA, Observation_Value)) %>% 
  filter(!is.na(diff_avg))

# calculate trimester

count_2 <- complete %>% group_by(Dim_Patient) %>% summarise(after_count = n())

count_rows <- merge(count_rows, count_2, by = "Dim_Patient")

low <- count_rows %>% filter(after_count <2)

view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#45556"))

(dim(df)[1] - dim(complete)[1]) / dim(df)[1] * 100

write.csv(weekly, "check.csv")

write.csv(df, "outliers_present.csv")

write.csv(count_rows, "count_rows.csv")

write.csv(low, "low_count.csv")




```



```{r}
# set the right baseline


initial_weight <- df %>% arrange(Dim_Patient, Observed_At) %>% group_by(Dim_Patient, GA_at_Observation) %>% filter(week == min(week)) %>% mutate(mode_wtt = getmode(Observation_Value))


write_csv(initial_weight, "initial_weight.csv")

view(initial_weight %>% group_by(Dim_Patient) %>% count() %>% filter(n>1))

```

Now, check if difference between the trimester is not greater than 15, remove those rows if greater than 15

```{r}

# Create the function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}


p1 = df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#1037")

remove_outliers <- function(df){

  print(dim(df))
  #check <- df %>% group_by(Dim_Patient, GA_at_Observation) %>% arrange(Observed_At) %>% mutate(mode = getmode(Observation_Value)) 
  
  
  check2 <- df %>% group_by(Dim_Patient) %>% arrange(Observed_At) %>% mutate(weight_diff = c(0,diff(Observation_Value)))
  
  check2 <- check2 %>% group_by(Dim_Patient, Trimester) %>% arrange(Observed_At) %>% mutate(weight_updated = ifelse( (abs(weight_diff) >= 15) &(Trimester != "PP"),NA, Observation_Value))
    
  check3 <- check2 %>% filter(!is.na(weight_updated))
  return(check3) 
  }

diff <- 1000
check <- df

p1 <- df %>% filter((Dim_Patient == "ref:Patient/BabyscriptsLegacy#4003") & (Observation_Value == 88.85))

df2 <- anti_join(df, p1, by = c("Dim_Patient","Observed_At"))

check <- df2 %>% group_by(Dim_Patient) %>% arrange(Observed_At) %>% mutate(weight_diff = c(0,diff(Observation_Value)))

while(diff >=15)
{
  check <- check %>% group_by(Dim_Patient, Trimester) %>% arrange(Observed_At) %>% mutate(weight_updated = ifelse((abs(weight_diff) >= 15) &(Trimester != "PP"),NA, Observation_Value))
    
  check <- check %>% filter(!is.na(weight_updated))
  
  check <- check %>% group_by(Dim_Patient) %>% arrange(Observed_At) %>% mutate(weight_diff = c(0,diff(Observation_Value)))

  sub <- check %>% filter(Trimester != "PP")
  diff <- max(abs(sub$weight_diff))
  print(diff)
}


length(unique(check$Dim_Patient))
length(unique(df$Dim_Patient))

dim(df) - dim(check)

write.csv(check, "cleaned_df.csv")

  
ct <- check %>% group_by(Dim_Patient) %>% count() %>% filter(n <3)


length(unique(check8$Dim_Patient))

ct <- check %>% filter((Trimester == "PP") & (weight_diff >25))

view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#1163"))

#df$weight_diff = ifelse(df$initial_weight == df$weight_diff, 0, df$weight_diff)

check <- df %>% group_by(Dim_Patient, Trimester) %>% filter(weight_diff > 15)



length(unique(df$Dim_Patient))

```



```{r}
initial_weight <- df %>% arrange(Dim_Patient, Observed_At) %>% group_by(Dim_Patient) %>%  slice(1) %>% select(Dim_Patient, Observation_Value) %>% rename(
    "initial_weight" = Observation_Value)
  
df <- merge(df, initial_weight,
              by.x= c("Dim_Patient"),
              by.y= c("Dim_Patient"),
              all.x =T)

func <- function(x) {
  
  w_diff <- c(0, diff(Observation_Value))>15
  
  weight_diff = ifelse(,0, weight_diff)
}





```


Remove if weeks are negative or greater than 50
```{r}

df$Trimester <- as.factor(df$Trimester)


# <0, >50, observations > 250
df <- df %>% filter((week >0) & (week < 50))
length(unique(df$Dim_Patient))

```


get min value of the observed at column for a patient
```{r}
# filter values, if weight is > 60% inital, (> %150 of final)
initial_weight <- df %>% arrange(Dim_Patient, Observed_At) %>% group_by(Dim_Patient) %>%  slice(1) %>% select(Dim_Patient, Observation_Value) %>% rename(
    "initial_weight" = Observation_Value)
  

#check4 <- dm %>% filter(Weight_Difference != 0.0)

#view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#2265"))

length(unique(check3$Dim_Patient))
  
df <- merge(df, initial_weight,
              by.x= c("Dim_Patient"),
              by.y= c("Dim_Patient"),
              all.x =T)



view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#1088"))

df <- df %>% filter(df$Observation_Value > 0.6*df$initial_weight.x)

```

```{r}
# check total rows of columns 
#check3 <- df %>% group_by(Dim_Patient) %>% count()
#sub2 <- check3 %>% filter(n > 250)
```

Group by patient id, trimester, week and take first entry
```{r}
df_week <- df %>% arrange(Dim_Patient, Observed_At) %>% group_by(Dim_Patient, week) %>%  slice(1)

check <-  df_week %>% group_by(Dim_Patient) %>% count()

#df_new <- df %>% group_by(Dim_Patient, Trimester, week) %>% summarise(mean_weight = #mean(Observation_Value))

```


Remove patients weight greater than 250
```{r}
#out <- df_week %>% filter(Observation_Value < 250)

length(unique(out$Dim_Patient))

# keep only week count greater than 5
check1 <- out %>% group_by(Dim_Patient, Trimester) %>% unique(Trimester)

out2 <-out %>% group_by(Dim_Patient, week) %>% count()

# calculate weight diff, check values 15

length(unique(out$Dim_Patient))
length(unique(out2$Dim_Patient))

```

## What is the change in weight during pregnancy for our entire population (Initial entry to 40weeks gestation) 
```{r}
q1 <-  out %>% group_by(week) %>% summarise(total_weight = sum(Observation_Value)) %>% arrange(week) %>% mutate(change = (c(total_weight[1], diff(total_weight)))/total_weight[1])

plot1 <- ggplot(data=q1, aes(x=week, y=change)) +
  geom_line()+
  geom_point() +
  ggtitle("Change in weight for all patients") 
  


ggplotly(plot1)
```


```{r}

# BMI between 15 and 65


# after removing less than 100 observ value, set first date as initial, calculate BMI  
# remove patients with height less than 1m 


# remove if weeks less than 2, 

check1 <- df %>% group_by(Dim_Patient, Trimester, week) %>% count()

view(df %>% filter(Dim_Patient == "ref:Patient/Redox#wellspan.org|007164527"))

#df$initial_weight <- ifelse(df$Days_Since_Last_Entry == FALSE, df$Observation_Value, 0)



```

## What is the change in weight for specific population groups (By customer group, by BMI category, by age) 


```{r}
# get only trimester values at patient-trimester level
dpt <- df %>% select(Dim_Patient, Trimester) %>% distinct(Dim_Patient, Trimester)
# check how many patients have length
agg <- aggregate(data=dpt, Trimester ~ Dim_Patient, function(x) length(unique(x)))
```

```{r}


# customers having full trimester data
df_tri <- agg %>% filter(Trimester == 3) %>% select(Dim_Patient)

#print(" Unique Patients having 3 trimester data", length(unique(df_tri$Dim_Patient)))

# filter to get entire dataset of these patients
sub_df <- merge(df_new, df_tri, by = "Dim_Patient", all.y = TRUE)
length(unique(sub_df$Dim_Patient))


# check week values for trimester patient data 
check <- sub_df %>% group_by(Dim_Patient) %>% summarise(min = min(week), max = max(week), count = n())


```

```{r}
check <- df %>% filter(Trimester != 'PP') %>% group_by(Dim_Patient) %>% count(GA_at_Observation)

```

```{r}
summary(df)
```

```{r}
sapply(df, function(x) sum(is.null(x)))

```


