---
title: "Patient data analysis"
author: "Tanaya Kavathekar"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    fig_height: 4.5
    fig_width: 7
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: true
  pdf_document:
    toc: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(error = F)
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = T)
knitr::opts_chunk$set(warning = F)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }


# Create the function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

```

```{r, include=F}
loadPkg("readxl")
loadPkg("plyr")
loadPkg("dplyr")
loadPkg("stringr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg("plotly")
loadPkg("lubridate")

```

### Reading data ..
```{r, include=F}
raw_df <- read.csv("GWG_5_18.csv")
names(raw_df) <- str_replace_all(names(raw_df), c(" " = "_"))

test_pt <- read_excel('test_patients_exclude.xlsx')
names(test_pt) <- str_replace_all(names(test_pt), c(" " = "_"))
length(unique(test_pt$Dim_Patient))

# if space in the column name change to underscore

#view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#3960"))
dim(raw_df)
length(unique(raw_df$Dim_Patient))
```

Total number of rows and columsn `r dim(raw_df)` and unique ids: `r length(unique(raw_df$Dim_Patient))`

### Removing test patients

already removed the count should not change, added as a testing step
```{r}
raw_df <- raw_df %>% filter(!Dim_Patient %in% test_pt$Dim_Patient)
```

Total test patients are `r length(unique(test_pt$Dim_Patient))` and `r dim(raw_df)`

### Drop null values from the data - 
```{r}
df <- raw_df %>% drop_na(Dim_Patient)

df$date <- as.Date(df$Observed_At)
df <- df %>% filter(Dim_Patient !=0)
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`

### Remove duplicate rows:
```{r}
# drop duplicates
df <- distinct(df, Dim_Patient, date, Observation_Value,  .keep_all= TRUE)
dim(df)
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


Create columns week, days and trimester
```{r}

df <- df %>% arrange(Dim_Patient, Observed_At)
df$EDD <- as.Date(df$EDD)


# add week and day column
df <- df %>% 
  mutate(week = trunc((279-as.numeric(EDD) + as.numeric(date))/7), 
         days = (279-as.numeric(EDD) + as.numeric(date))%%7, 
         trimester = ifelse(trunc((279-as.numeric(EDD) + as.numeric(date))/7)< 13,"T1",
                                                          ifelse(trunc((279-as.numeric(EDD) + 
                                                                          as.numeric(date))/7)< 27, "T2",
                                                                  ifelse(trunc((279-as.numeric(EDD) 
                                                                                + as.numeric(date))/7)<41,"T3","PP"))))

```



### Drop duplicated rows which have multiple enteries on the same day
```{r}
#df <-subset(df,select = - c(Observed_At, week, days, trimester, weight_diff, wt_flg))
df <-subset(df,select = - c(week, days, trimester, X))

df <- unique(df)

dim(df)
length(unique(df$Dim_Patient))
```


### Removing rows having null in EDD column
```{r, include=F}
# GTA column has missing values 
sum(is.na(df))
# filter rows having EDD as null
gat_null <- df %>% filter(is.na(EDD))

# time being drop those columns 
df <-df %>% drop_na(EDD)
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


### Filter patients between 80 and 380 observation values
```{r , include=T}
df <- df %>% filter((Observation_Value > 80) & (Observation_Value < 380))
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


### Keep only 47 weeks of data from observe date
```{r, include=T}
df$EDD  <- as.Date(df$EDD)
# add week and day column
df <- df %>% 
  mutate(week = trunc((279-as.numeric(EDD) + as.numeric(date))/7), 
         days = (279-as.numeric(EDD) + as.numeric(date))%%7, 
         trimester = ifelse(trunc((279-as.numeric(EDD) + as.numeric(date))/7)< 13,"T1",
                                                          ifelse(trunc((279-as.numeric(EDD) + 
                                                                          as.numeric(date))/7)< 27, "T2",
                                                                  ifelse(trunc((279-as.numeric(EDD) 
                                                                                + as.numeric(date))/7)<41,"T3","PP"))))

df <- df %>% filter((week>=0) &(week<=47))

length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


### Remove patients having only less than 4 observations and only one trimester data 
```{r, include=F}
# get only trimester values at patient-trimester level
dpt <- df %>% select(Dim_Patient, trimester) %>% distinct(Dim_Patient, trimester)

# check how many patients have length
agg <- aggregate(data=dpt, trimester ~ Dim_Patient, function(x) length(unique(x)))

less_data <- agg %>% filter(trimester == 1)

length(unique(less_data$Dim_Patient))

check2 <- merge(
  df, less_data, by = "Dim_Patient", all.y=T)


check2_2 <- check2 %>% group_by(Dim_Patient) %>% count()

# remove these patients 
ct2 <- check2_2 %>% filter(n<4)
length(unique(ct2$Dim_Patient))


df <- subset(df, !df$Dim_Patient %in% ct2$Dim_Patient)
length(unique(df$Dim_Patient))

# save data of removed patients 
#abc <- check2 %>% filter(Dim_Patient %in% ct2$Dim_Patient) #12789

#write.csv(abc, "less_entries_5_18.csv")
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


### Remove entries if less than 2 observations within a day
```{r}
intermediate_rollup <- df %>% group_by(Dim_Patient, date) %>% summarise(mean = mean(Observation_Value))

count <- intermediate_rollup %>% group_by(Dim_Patient)  %>% count() %>% filter(n > 2)

df <- merge(df, count[,"Dim_Patient"], by = "Dim_Patient")
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


Calculate trimester and days since column
```{r}
df <- df %>% group_by(Dim_Patient) %>% 
  arrange(Dim_Patient, Observed_At) %>%
  mutate(days_since = c(0,diff(date)))
# view(df %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#1177"))
```


Keep the manually cleaned data
```{r}
man_cl1 <- read_xlsx('low_count.xlsx')
keep1 <- man_cl1 %>% filter((Resolution == "KEEP")| (Resolution == "keep"))
remove1 <- man_cl1 %>% filter((Resolution == "REMOVE")| (Resolution == "remove"))

man_cl2 <- read_xlsx('low_count2.xlsx')
keep2 <- man_cl2 %>% filter((Resolution == "KEEP") | (Resolution == "keep"))
remove2 <- man_cl2 %>% filter((Resolution == "REMOVE")| (Resolution == "remove"))


man_cl3 <- read_xlsx('low_weight_patients.xlsx')
man_cl3 <- man_cl3 %>% filter(!(Dim_Patient %in% c("ref:Patient/Athena#13122|2690580", 
                              "ref:Patient/BabyscriptsLegacy#6451", 
                              "ref:Patient/Redox#advocateaurorahealth.org|1699226", 
                              "ref:Patient/Redox#advocateaurorahealth.org|2583557", 
                              "ref:Patient/Redox#advocateaurorahealth.org|4394288")))

high_dev <- read_excel('high_deviation_corrected.xlsx')
high_dev <- unique(high_dev)
high_dev <- high_dev  %>% 
  select(Dim_Patient, Observation_Source, Observation_Value, EDD, GA_at_Observation, date) %>% rename("Observed_At" = date)


manually_rm_out <- df %>% filter((Dim_Patient %in% keep1$Dim_Patient) | 
                                   (Dim_Patient %in% keep2$Dim_Patient) | 
                                   (Dim_Patient %in% man_cl3$Dim_Patient) |
                                   (Dim_Patient %in% high_dev$Dim_Patient) 
                                   #(Dim_Patient %in% low_count3$Dim_Patient)
                                 )

df <- df %>% 
  filter(!(Dim_Patient %in% remove1$Dim_Patient)) %>% 
  filter(!(Dim_Patient %in% remove2$Dim_Patient)) %>% 
  filter(!(Dim_Patient %in% c("ref:Patient/BabyscriptsLegacy#41617", 
                              "ref:Patient/BabyscriptsLegacy#49246", 
                              "ref:Patient/BabyscriptsLegacy#5622", 
                              "ref:Patient/Redox#advocateaurorahealth.org|912816")))
  

length(unique(manually_rm_out$Dim_Patient))
```


### Day level and week level logic
```{r}

temp_df <- df %>% filter(!(Dim_Patient %in% manually_rm_out$Dim_Patient)) 
#%>% filter(!Dim_Patient %in% low2$Dim_Patient)


# count initial values
count_rows <- temp_df %>% group_by(Dim_Patient, date) %>%  
  summarise(wt = mean(Observation_Value)) %>% ungroup() %>%  
  group_by(Dim_Patient) %>% summarise(initial_count = n())

daily <- temp_df %>% group_by(Dim_Patient, date) %>% arrange(Dim_Patient, Observed_At) %>%
  mutate(weight_diff = c(0,diff(Observation_Value)), 
         wt_flg =  ifelse((abs(weight_diff) >= 5) & (trimester != "PP"), NA, Observation_Value))  %>% ungroup() %>% filter(!is.na(wt_flg))

weekly2 <- daily %>% group_by(Dim_Patient, trimester) %>% arrange(Dim_Patient, Observed_At) %>%
  mutate(weight_diff_tri = c(0,diff(Observation_Value)), 
         mode_wt_tri = getmode(Observation_Value),
         diff = Observation_Value - mode_wt_tri,
         wt_flg_tri =  ifelse((days_since <= 70) & (abs(diff) >= 25) & (trimester != "PP"), NA, Observation_Value)) %>%
  filter(!is.na(wt_flg_tri))%>% 
  ungroup() %>% 
  group_by(Dim_Patient) %>% 
  mutate(mode = getmode(Observation_Value), mode_dev =  ifelse((abs(Observation_Value - mode) >= 40), NA, Observation_Value)) %>% 
  filter(!is.na(mode_dev))


complete <- weekly2 %>% group_by(Dim_Patient, trimester) %>% 
  mutate( avg = mean(Observation_Value)) %>% 
  ungroup() %>% 
  group_by(Dim_Patient) %>% 
  mutate(diff_avg =ifelse((abs(Observation_Value - avg[1]) >= 20) & (trimester == "PP"), NA, Observation_Value)) %>% 
  filter(!is.na(diff_avg))

# calculate filtered counts
count_2 <- complete %>% group_by(Dim_Patient, date) %>%  
  summarise(wt = mean(Observation_Value)) %>% ungroup() %>%  
  group_by(Dim_Patient) %>% summarise(after_count = n())

count_rows <- merge(count_rows, count_2, by = "Dim_Patient") 

count_rows <- count_rows %>% mutate(per_removed = ((initial_count - after_count)/initial_count) * 100)
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`



Remove low count and add manually filtered data back to the cleaned df
```{r}
low2 <- count_rows %>% filter(per_removed  > 50)

# low_new <- count_rows %>% filter(per_removed  > 50)
# 
# fix_this <- df %>% filter(Dim_Patient %in% low$Dim_Patient)
# write.csv(fix_this, "fix_this.csv")

# view(complete %>% filter(Dim_Patient == "ref:Patient/BabyscriptsLegacy#16707"))

cleaned_df <- complete %>% 
# %>% filter(!(Dim_Patient %in% low$Dim_Patient))
 select(Dim_Patient, Observation_Source, Observation_Value, Observed_At,
        EDD, GA_at_Observation, date, week, trimester,days, days_since)


processed_df <- rbind(cleaned_df, manually_rm_out)
length(unique(processed_df$Dim_Patient))
summary(processed_df)

```

Total number of rows and columns are `r dim(processed_df)` and total number of unique patients are `r length(unique(processed_df$Dim_Patient))`

```{r}
#write.csv(processed_df, "processed_df_5_19.csv", row.names=FALSE)
sum(is.na(processed_df$days))
```





```{r}
# check T3 drop  wt diff if < -9
```