---
title: "Data Merge"
author: "Tanaya Kavathekar"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    fig_height: 4.5
    fig_width: 7
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: true
  pdf_document:
    toc: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(error = F)
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = T)
knitr::opts_chunk$set(warning = F)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }


# Create the function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

```

```{r, include=F}
loadPkg("readxl")
loadPkg("plyr")
loadPkg("dplyr")
loadPkg("stringr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg("plotly")
loadPkg("lubridate")

```

##### Reading data ..
```{r, include=F}
raw_df <- read_excel('GWG(2+)_Corrected.5.10.20(lowcount).xlsx',sheet='master')
names(raw_df) <- str_replace_all(names(raw_df), c(" " = "_"))
length(unique(raw_df$Dim_Patient))

test_pt <- read_excel('test_patients_exclude.xlsx')
names(test_pt) <- str_replace_all(names(test_pt), c(" " = "_"))
test_pt <- test_pt %>% select(Dim_Patient) %>% unique()
length(unique(test_pt$Dim_Patient))

```
Total number of rows and columsn `r dim(raw_df)` and unique ids: `r length(unique(raw_df$Dim_Patient))`


Removed test patients from the raw data. Not total number of rows and columns are `r dim(raw_df)` and total number of unique patients are `r length(unique(raw_df$Dim_Patient))`

##### Drop null values from the data - 
```{r, include=FALSE}
df <- raw_df %>% drop_na(Dim_Patient)
df <- df %>% filter(Dim_Patient !=0)
# remove diff column as it will be recalculated again
df <-subset(df,select = - c(Weight_Difference, Days_Since_Last_Entry, Trimester))
dim(df)
length(unique(df$Dim_Patient))


```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


Merging high deviation data
```{r}
high_dev <- read_excel('high_deviation_corrected.xlsx')
high_dev <- unique(high_dev)
high_dev$EDD <- high_dev$EDD

high_dev <- high_dev  %>% 
  select(Dim_Patient, Observation_Source, Observation_Value, EDD, GA_at_Observation, date) %>% rename("Observed_At" = date)


high_dev_rm <- df %>% filter(!Dim_Patient %in% high_dev$Dim_Patient)


df <- rbind(high_dev_rm, high_dev)
dim(df)
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`


Merging low count 1
```{r}
# merge corrected low count df
low_count <- read_excel('original_df.xlsx')
#low_count$date <- as.Date(low_count$Observed_At)
low_count <-subset(low_count,select = - c(Trimester))

df_low_count_rm <- df %>% filter(!Dim_Patient %in% low_count$Dim_Patient)

df <- rbind(df_low_count_rm, low_count)
dim(df)
length(unique(df$Dim_Patient))
```
Total number of rows and columns are `r dim(df)` and total number of unique patients are `r length(unique(df$Dim_Patient))`

Merging low count 2
```{r}
low_count2 <- read_excel('original_df2_updated.xlsx')

low_count2 <- low_count2  %>% 
  select(Dim_Patient, Observation_Source, Observation_Value, EDD, GA_at_Observation, date) %>% rename("Observed_At" = date)


df_low_count_rm2 <- df %>% 
  filter(!Dim_Patient %in% low_count2$Dim_Patient)

df <- rbind(df_low_count_rm2, low_count2)
dim(df)
length(unique(df$Dim_Patient))
```

##### Removing test patients 
```{r, include=F}
df <- df %>% filter(!Dim_Patient %in% test_pt$Dim_Patient)
length(unique(df$Dim_Patient))

df_c <- unique(df)

```
Total test patients are `r length(unique(test_pt$Dim_Patient))`

```{r}
write.csv(df, "GWG_5_18.csv")
```

